Dispersion distance and the matter distribution of the Universe in dispersion space

### 2015-06-04 writeup/dm_space_distortions.tex
Many small edits, mostly to makes things slightly more breif.

\documentclass[twocolumn,prl,nofootinbib,floatfix]{revtex4-1}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\usepackage{verbatim}
\usepackage{times}
\usepackage{afterpage}
\usepackage{epstopdf}

@@

However, redshift is not the only measure of radial distance. Standard
candles and standard rulers can be used to estimate distance using brightness
and angular size respectively.  In the same way a \emph{standard ping}, a brief
broadband radio impulse, can be used to measure dispersion through the
characteristic wavelength-squared delay of electromagnetic signals. The disperion is
proportional to the amount of plasma along the
line-of-sight and can thus be used to estimate radial distance.
This technique, routinely done with
pulsars in the Galaxy, can be used to define a \emph{dispersion distance} to
cosmological sources.
%
Dispersion is readily measured from fast radio bursts (FRB),
a new class of millisecond duration radio transients first reported by
\citet{2007Sci...318..777L}, which are the archetype of a standard ping.
Ten such bursts have been reported to date
\citep{2012MNRAS.425L..71K,2013Sci...341...53T, 2014ApJ...790..101S,
2014ApJ...792...19B, 2015MNRAS.447..246P,2015ApJ...799L...5R}, and the total
event rate has been estimated to be $3.3^{+5}_{-2.5} \times 10^3$ per sky per
day \cite{2015arXiv150500834R}.
Dispersion is characterized by a line-of-site
integral of the free electron density known the dispersion measure (DM).  FRB
events are observed to have dispersion measures of order $\sim\!\!1000\,{\rm pc}
/ {\rm cm}^3$, greatly in excess of the galactic expectation of order
$\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on galactic latitude)
\cite{2014arXiv1412.4829D}.  They thus appear to be of extra-galactic origin,
with their dispersion produced either by the inter-galactic medium (IGM)
surrounding galaxies along the line-of-sight \citep{2013Sci...341...53T}
or the environment of the
source \citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.  If
the dispersion is dominated by electrons along the line-of-sight, then the
average dispersion--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I} and the
sources appear to be at cosmological distances of of order $\sim\!\!{\rm Gpc}$.
Such cosmological distances permit the study of clustering of FRB sources in
three dimensions with an expectation that these events trace the large-scale
structure of the Universe on linear scales. We note that while our discussion
here is framed in terms of FRBs, other bright radio transients that fluctuate
on timescales less than the typical differential lag of $\sim 1s/100 \,{\rm
MHz} $ might also be useful as standard pings.

@@

The dispersion measure of a signal observed in some angular direction $\hat{n}$ and
originating from comoving radial distance $\chi$ is
\begin{equation}
    {\rm DM}(\hat{n}, \chi) = \int_0^{\chi} \ud\chi' a(\chi)^2
        n_e(\hat{n}\chi', \chi').
\end{equation}
Here
$n_e(\vec{x}, \chi)$ is the free electron density as a function of location and
conformal time. Note that we use $\chi$ as our radial distance and time
coordinate, as opposed to redshift which is more appropriate for spectroscopic
applications. We model the cosmological 
electron density as containing a homogeneous part
and perturbations,
$n_e(\vec x, \chi) = \bar{n}_e(\chi) \left[1 + \delta_e(\vec x, \chi)\right]$.
The dispersion
measure is thus
\begin{equation}
    {\rm DM}(\hat n, \chi) = \int_0^\chi \ud\chi' a(\chi')^2 \bar{n}_e(\chi')
       \left[1 + \delta_e(\hat{n}\chi',\chi')\right].
\end{equation}

@@

The Jacobian in spherical coordinates is
\begin{align}
\left| \frac{\ud^3\vec x}{\ud^3\vec x_s} \right|
    &= \frac{\ud \chi}{\ud \chi_s}\frac{\chi^2}{\chi_s^2}\\
    &= (1 + \delta_e)^{-1}
       \left(1 + \frac{\int_0^\chi \ud \chi' \delta_e(\hat n \chi')}
                      {\chi}\right)^{-2}\\
    &\approx 1 - \delta_e
        - \frac{2}{\chi}\int_0^\chi \ud \chi' \delta_e(\hat n \chi').
        \label{e:jac}
\end{align}

Substituting Equations~\ref{e:nfs}~and~\ref{e:jac} into Equation~\ref{e:density},
we obtain
\begin{equation}
\label{e:delta_s}
    \delta_{fs} = \delta_f - \delta_e
    - \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
        \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}
In the equation above, the $-\delta_e$ term is most analogous to the Kaiser
redshift-space distortions.  It is a dilution of tracers in dispersion space due to an
excess of electrons between the tracers. However we note that this term is isotropic
in contrast to the Kaiser term.  This is because any wave vector electron perturbation
causes dispertion space distortions, whereas the radial velocities that cause
redshift-space distortions are only sourced by perturbations with radial wave
vectors.

@@

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike the redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the dispersion-space over density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependence, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power spectrum is negligible over these time
differences which leads to a straight-forward way to include the evolution of
the perturbations:
\begin{align}
C^{xy}_\ell(\chi,\chi') 
    \approx \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,(\chi + \chi')/2),\nonumber &\\
    |\chi - \chi'| \ll 1/aH.&
\end{align}

@@

The second term in Equation~\ref{e:Clss_limber} is an integral that mixes
many spatial scales: small scales from smaller radial distance and larger
scales from greater distances. The mixing of scales complicates the
interpretation of measurements, since structure formation on scales smaller
than $\sim10\,{\rm Mpc}/h$ is non-linear and hard to model. A similar issue
exists in the field of weak gravitational lensing, where the formula for the
angular power spectrum has a similar form.  In lensing, a kernel in
the line-of-sight integral arising from geometric effects suppresses the
contribution from small distances, partially alleviating the issue.

There is an advantage to the lack of kernel in the dispersion-space integral term.
Lensing tomography---the use of sources at multiple redshifts to un-mix
spatial scales---is complicated by the kernel. The kernel's shape depends on
the source redshift, re-weighting the line-of-sight integral and
making tomography inexact.  In contrast, the integrand in the dispersion-space integral
term is completely independent of the limit of integration, meaning
contributions from different parts of the line of sight (and thus different
spatial scales) can be separated exactly. A detailed study of dispersion space
tomography is beyond the scope of this work.  To avoid our results being
sensitive to the hard-to-model small scales, we assume that the contribution to the
integral term from radial distances $\chi'' < 500\,{\rm Mpc}/h$ are well
measured by the correlations with these radial slices. We thus ignore
contributions to the integral term from below this distance and neglect the
information from these slices in our sensitivity measurements. The smallest
spatial scale that contributes the our plotted angular power spectra out to
$\ell=1000$ is thus $k=2\,h/{\rm Mpc}$.

@@

There are a number of effects that will complicate the measurement of the
above calculated angular power spectrum by a real survey of dispersion 
measures.  The first effect is
that the model that the free electron density decomposes into a homogeneous
part plus a random field of perturbations is not complete. Firstly, all signals
must pass through the Milky Way's galactic disk, halo, and local environment,
which will contribute to the dispersion. However, this is only a
single function of $\hat n$ which can presumably be well
measured. In addition, this function could in principle be well characterized
by independent data such as observations of pulsars.  As such, it should be
possible to subtract the local
contribution, as was done in
\citet{2014arXiv1412.4829D}, and this is unlikely to be a limiting obstacle.

Perhapse, more concerning is the dispersion from the source's environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than dispersion-space
structure map pixels and thus are not accounted for in the $\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean dispersion--distance relation. This will need to be calibrated in some
why or else fit with nuisance parameters during parameter estimation.  There
will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of dispersion space density maps in the radial direction.

@@

In first measurements of the angular power spectra, uncertainty will be
dominated by the finite number of observed events, or shot noise.
The noise power spectrum on shells with of width
$\Delta \chi$ is
\begin{equation}
    C^N_{\ell, ij} = \frac{\delta_{ij}}{\bar n_f(\chi_i)\chi_i^2\Delta\chi}.
\end{equation}
Here, to indicate that we are now working with quantities binned in the radial
direction instead of
continuous functions, we've switched to $C_{\ell,ij}$ instead of
$C_\ell(\chi, \chi')$. The uncertainty on the angular
cross-power spectrum is then
\begin{align}
    (\Delta{C^{ss}_{\ell,ij}})^2 = 
    {\frac{1}{\ell(\ell + 1)\Delta \ell f_{\rm sky}}}
        \left[C^N_{\ell,ii} C^N_{\ell,jj} +  (C^N_{\ell,ij})^2\right],
\end{align}
where the second term in brackets only contributes for $i = j$.

We consider the sensitivity of a survey with a total of $10000$ dispersion
measures from FRB events distributed uniformly over half the sky. This is in rough
approximation to what might be achieved by the Canadian Hydrogen Intensity
Mapping Experiment (CHIME) telescope over a three year survey.
To give an idea of the overall sensitivity of the survey, we collapse
$C^{ss}_{\ell,ij}$ to a single function of $\ell$ by taking a weighted average
of all pairs of radial bins over the
range $500\,{\rm Mpc}/h$ to $3500\,{\rm Mpc}/h$.
For weights we use the signal over noise-squared
at $\ell=100$, which maximizes the signal to noise at that $\ell$ and
is near optimal at other multipoles.
The result is shown
in Figure~\ref{f:sensitivity}.

@@

We have proposed using FRB sources as 3D tracers of the large-scale structure,
using dispersion to estimate radial distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the line of sight inducing mis-estimates of radial
distance and thus apparent clustering. These dispersion-space distortions could be used
to study the distribution of free electrons in the Universe on large-scales and
redshifts less than $\sim\!\!1$.
There have been other proposals to use FRB events to study large-scale
structure \citep{2014ApJ...780L..33M,2014MNRAS.443L..11D}, however these
schemes have either been two dimensional or required externally measured redshifts
for the sources. Additionally, there have been proposals to use the dispersion-redshift
relation to probe the background density of electrons and thus the expansion
history of the Universe
\citep{2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}.

We have
shown that a survey detecting $10000$ fast radio bursts of cosmological
origin (as might be achieved by the CHIME FRB instrument) could detect the
clustering signal.
Looking forward, dispersion-space
clustering could become a powerful probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion from
the source's environment and host halo be small compared to the
contribution from the IGM.

>>>>

\documentclass[twocolumn,prl,floatfix]{revtex4-1}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\usepackage{verbatim}
\usepackage{times}
\usepackage{afterpage}
\usepackage{epstopdf}
\usepackage{url}
\usepackage{hyperref}

@@

However, redshift is not the only measure of radial distance. Standard
candles and standard rulers can be used to estimate distance using brightness
and angular size respectively.  In the same way a \emph{standard ping}, a brief
broadband radio impulse, can be used to measure dispersion through the
characteristic wavelength-squared delay of electromagnetic signals. The dispersion is
proportional to the amount of plasma along the
line-of-sight and can thus be used to estimate radial distance.
This technique, routinely used for
pulsars in the Galaxy, can be used to define a \emph{dispersion distance} to
cosmological sources.
%
Dispersion is readily measured from fast radio bursts (FRB),
a new class of millisecond duration radio transients first reported by
\citet{2007Sci...318..777L}, which are the archetype of a standard ping.
Ten such bursts have been reported to date
\citep{2012MNRAS.425L..71K,2013Sci...341...53T, 2014ApJ...790..101S,
2014ApJ...792...19B, 2015MNRAS.447..246P,2015ApJ...799L...5R}, and the total
event rate has been estimated to be $3.3^{+5}_{-2.5} \times 10^3$ per sky per
day \cite{2015arXiv150500834R}.
Dispersion is characterized by a line-of-site
integral of the free electron density known the dispersion measure (DM).  FRB
events are observed to have dispersion measures of order $\sim\!\!1000\,{\rm pc}
/ {\rm cm}^3$, greatly in excess of the galactic expectation of order
$\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on galactic latitude)
\cite{2014arXiv1412.4829D}.  They thus appear to be of extra-galactic origin,
with their dispersion produced either by the inter-galactic medium (IGM),
associated with structure along line-of-sight \citep{2013Sci...341...53T},
or the environment of the
source \citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.  If
the dispersion is dominated by electrons along the line-of-sight, then the
average dispersion--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I} and the
sources appear to be at cosmological distances of of order $\sim\!\!{\rm Gpc}$.
Such cosmological distances permit the study of clustering of FRB sources in
three dimensions with an expectation that these events trace the large-scale
structure of the Universe on linear scales. We note that while our discussion
here is framed in terms of FRBs, other bright radio transients that fluctuate
on timescales shorter than the typical differential lag of $\sim 1s/100 \,{\rm
MHz} $ might also be useful as standard pings.

@@

The dispersion measure of a signal observed in some angular direction $\hat{n}$ and
originating from comoving radial distance $\chi$ is
\begin{equation}
    {\rm DM}(\hat{n}, \chi) = \int_0^{\chi} \ud\chi' a(\chi)^2
        n_e(\hat{n}\chi', \chi').
\end{equation}
Here
$n_e(\vec{x}, \chi)$ is the free electron density as a function of location and
conformal time. Note that we use $\chi$ as our radial distance and time
coordinate, as opposed to redshift. We model the cosmological 
electron density as containing a homogeneous part
and perturbations,
$n_e(\vec x, \chi) = \bar{n}_e(\chi) \left[1 + \delta_e(\vec x, \chi)\right]$.
The dispersion
measure is thus
\begin{equation}
    {\rm DM}(\hat n, \chi) = \int_0^\chi \ud\chi' a(\chi')^2 \bar{n}_e(\chi')
       \left[1 + \delta_e(\hat{n}\chi',\chi')\right].
\end{equation}

@@

The Jacobian in spherical coordinates is
\begin{align}
\left| \frac{\ud^3\vec x}{\ud^3\vec x_s} \right|
    &= \frac{\ud \chi}{\ud \chi_s}\frac{\chi^2}{\chi_s^2}\\
%    &= (1 + \delta_e)^{-1}
%       \left(1 + \frac{\int_0^\chi \ud \chi' \delta_e(\hat n \chi')}
%                      {\chi}\right)^{-2}\\
    &\approx 1 - \delta_e
        - \frac{2}{\chi}\int_0^\chi \ud \chi' \delta_e(\hat n \chi').
        \label{e:jac}
\end{align}

Substituting Equations~\ref{e:nfs}~and~\ref{e:jac} into Equation~\ref{e:density},
we obtain
\begin{equation}
\label{e:delta_s}
    \delta_{fs} = \delta_f - \delta_e
    - \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
        \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}
In the above equation, the $-\delta_e$ term is most analogous to the Kaiser
redshift-space distortions.  It is a dilution of tracers in dispersion space due to an
excess of electrons between the tracers. However we note that this term is isotropic
in contrast to the Kaiser term.  This is because any wave vector electron perturbation
causes dispersion space distortions, whereas the radial velocities that cause
redshift-space distortions are only sourced by perturbations with radial wave
vectors.

@@

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike the redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the dispersion-space over density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependence, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power spectrum is negligible over these time
differences which leads to a straight-forward way to include the evolution of
the perturbations:
\begin{align}
    % I've tuned the alignment of the domain (using &)
    % here to get the equation number and
    % domain on the same line. Not elegant.
    C^{xy}_\ell(\chi,\chi') \approx&
        \frac{2}{\pi}
        \int_0^\infty\ud k k^2
        j_\ell(k\chi) j_{\ell}(k\chi')
        P_{xy}(k,(\chi + \chi')/2),\nonumber \\
    &|\chi - \chi'| \ll 1/aH.
\end{align}

@@

The second term in Equation~\ref{e:Clss_limber} is an integral that mixes
many spatial scales: small scales from smaller radial distance and larger
scales from greater distances. The mixing of scales complicates the
interpretation of measurements, since structure formation on scales smaller
than $\sim10\,{\rm Mpc}/h$ is non-linear and hard to model. A similar issue
exists in the field of weak gravitational lensing, where the formula for the
shear angular power spectrum has a similar form.  In lensing, a kernel in
the line-of-sight integral arising from geometric effects suppresses the
contribution from small distances, partially alleviating the issue.

However, the lack of a kernel simplifies tomography---the use of sources at
multiple redshifts to unmix spatial scales.
Lensing tomography is inexact because the kernel's shape depends on
the source redshift, thus re-weighting the line-of-sight integral.
In contrast, the integrand in the dispersion-space integral
term is completely independent of the limit of integration, meaning
contributions from different parts of the line of sight (and thus different
spatial scales) can be separated exactly. A detailed study of dispersion space
tomography is beyond the scope of this work.  To avoid our results being
sensitive to the hard-to-model small scales, we assume that the contribution to the
integral term from radial distances $\chi'' < 500\,{\rm Mpc}/h$ are well
measured by the correlations with these radial slices. We thus ignore
contributions to the integral term from below this distance and neglect the
information from these slices in our sensitivity measurements. The smallest
spatial scale that contributes to our plotted angular power spectra out to
$\ell=1000$ is thus $k=2\,h/{\rm Mpc}$.

@@

In a real survey, measurement of the dispersion-space clustering
will be complicated by extra contributions to the dispersion.
Firstly, all signals
must pass through the Milky Way's galactic disk, halo, and local environment.
However, the resulting dispersion is only a
single function of $\hat n$ which can presumably be well
measured. As such, it should be
possible to subtract the local
contribution, as was done in
\citet{2014arXiv1412.4829D}, and this is unlikely to be a limiting obstacle.

Perhaps, more concerning is the dispersion from the source's immediate environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than the survey
resolution and are thus not accounted for in the
$\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean dispersion--distance relation. This will need to be calibrated in some
way or else fit with nuisance parameters during parameter estimation.  There
will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of dispersion space density maps in the radial direction.

@@

In first measurements of the angular power spectra, uncertainty will be
dominated by the finite number of observed events, or shot noise.
The noise power spectrum on shells at distance with of width
$\Delta \chi$ is
\begin{equation}
    C^N_{\ell, ij} = \frac{\delta_{ij}}{\bar n_f(\chi_i)\chi_i^2\Delta\chi}.
\end{equation}
Here, to indicate that we are now working with quantities binned in the radial
direction, we've switched to $C_{\ell,ij}$ instead of
$C_\ell(\chi, \chi')$. The uncertainty on the angular
cross-power spectrum is then
\begin{align}
    (\Delta{C^{ss}_{\ell,ij}})^2 = 
    {\frac{1}{\ell(\ell + 1)\Delta \ell f_{\rm sky}}}
        \left[C^N_{\ell,ii} C^N_{\ell,jj} +  (C^N_{\ell,ij})^2\right],
\end{align}
where the second term in brackets only contributes for $i = j$.

We consider the sensitivity of a survey with a total of $10000$ dispersion
measures from FRB events distributed uniformly over half the sky.
To give an idea of the overall sensitivity of the survey, we collapse
$C^{ss}_{\ell,ij}$ to a single function of $\ell$ by taking a weighted average
of all pairs of radial bins over the
range $500\,{\rm Mpc}/h$ to $3500\,{\rm Mpc}/h$.
For weights we use the signal over noise-squared
at $\ell=100$, which maximizes the signal to noise at that $\ell$ and
is near optimal at other multipoles.
The result is shown
in Figure~\ref{f:sensitivity}.

@@

We have proposed using cosmological FRB sources as 3D tracers of the
large-scale structure,
using dispersion to estimate radial distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the line of sight, inducing mis-estimates of radial
distance and thus apparent clustering. These dispersion-space distortions could
be used
to study the distribution of free electrons in the Universe on large-scales.
There have been other proposals to use FRBs in cosmology 
\citep{2014ApJ...780L..33M,2014MNRAS.443L..11D,
2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}, however these
schemes have either not been three dimensional
or required externally measured redshifts
for the sources.

We have
shown in Figure~\ref{f:sensitivity}
that a survey detecting $10000$ fast radio bursts of cosmological
origin could detect the
clustering signal. This is in rough
approximation to what might be achieved by the Canadian Hydrogen Intensity
Mapping Experiment (CHIME)\footnote{\url{http://chime.phas.ubc.ca/}}
or other upcoming wide-field telescopes.
Looking forward, dispersion-space
clustering could become a precision probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion from
the source's environment and host halo be small compared to the
contribution from the IGM.
